Synchrotron X-Ray Rims in Tycho's Supernova Remnant are Energy Dependent

### 2014-09-22 manuscript/paper-tycho.tex
Paper text, full model simulation, more tables
Tables for Tycho regions-5, notebook to check model fit sensitivity
to eta2. More introductory text in manuscript.

% This paper
\newcommand*{\tsup}{\textsuperscript}
\newcommand*{\Chandra}{\textit{Chandra}\ }
\newcommand*\mean[1]{\bar{#1}}
\newcommand*{\tsynch}{\tau_{\mt{synch}}}
\newcommand*{\mE}{m_\mt{E}}
\newcommand*{\Ecut}{E_{\mt{cut}}}
\defcitealias{ressler2014}{R14}

@@

% Intro -- shocks, DSA, thin rims, filament observations; other fields/work
Forward shock accelerated ISM/CSM particles in young supernova remnants emit
synchrotron radiation strongly in the shock's immediate wake but quickly turn
off downstream \citep{koyama1995, reynolds1996}, producing a shell-like
morphology with bright X-ray and radio rims/filaments due to line-of-sight
projection (hereafter, thin rims).
Seminal observations \& interpretations of these filaments, interpretation in
context of DSA.  Attempts to model filaments and figure out what's going on.
Why do we care about these models?  They tell us something about what happens
behind the shock, and perhaps what's going on at the shock.  We get at the
plasma and magnetic field turbulence, electron injection.
Applications to: cosmic ray acceleration, turbulent field amplification, (note
assumptions in re proton/electron spectra).
Cut-off energy of injected electrons?
Relevant to, broadly, all astrophysical shocks: Earth's bow shock,
PWNe, SNe and/or stellar wind blown bubbles, starbursts \citep{heckman1990},
AGN jets, galaxy clusters \citep{van-weeren2010}, and even cosmological shocks
\citep{miniati2000, ryu2008}.  See relevant reviews \citep{blandford1987}.

% What makes these shocks?
The widths of these X-ray rims are thought to be controlled by a combination of
synchrotron losses, particle transport, and magnetic fields immediately
downstream of the shock.  Synchrotron losses depend on the initial electron
energy distribution and the gradual decrease of electron energies downstream of
the shock.  High energy electrons near the shock efficiently radiate harder
synchrotron X-ray photons, but decline in energy as they are transported away
from the shock; this may be seen in spectral softening downstream of the shock
\citep[e.g.,][]{cassam-chenai2007}.  The transport process is driven by bulk
plasma advection downstream of the shock and diffusion (\emph{driven by
collisionless particle interaction, magnetic field turbulence, ?????  fill in
details}) with respect to bulk advection.  Energy-dependent diffusion (e.g.,
diffusion coefficient $D \propto E$ for Bohm diffusion) allows higher energy
electrons to diffuse further upstream or downstream than would be expected from
pure advection; in particular, diffusion upstream of the shock may give rise to
a shock precursor \citep[e.g.,][]{ghavamian2000, wagner2009, laming2014}.
Finally, the magnetic field strength may be quickly damped behind the shock, so
that electrons may not radiate as efficiently and thin rims would reflect the
spatial structure of the field, rather than efficient particle acceleration
(and losses) and/or efficient synchrotron cooling \citep{pohl2005}.

% Energy dependence.
Moreover, these controlling mechanisms predict different scalings for rim width
as a function of energy.  Synchrotron loss-limited rims should scale as $w
\propto \nu^{-1/2}$.  Diffusion smears out rims at all energies with stronger
effect at higher energies, yielding a slower drop-off in rim widths when
diffusion dominates.  Magnetic damping, if at a length scale comparable to
filament widths, predicts energy-independent rim widths -- intuitively, if the
magnetic field turns off, synchrotron radiation turns off regardless of
electron energy.  \citet{ressler2014} give a much fuller exposition of the
relative effects of advection, diffusion, and magnetic field damping.

% Which one is right?
Which mechanisms are most relevant to historical SNRs?
\note[Aaron]{need to read more / more thoroughly}
Introduce $\tsynch$ and the basic intuition. Explain why diffusion is important
(when did diffusion start being recognized as relevant to SNR filaments? or how was
this shown?).
Following Sean (i.e., taking all of his references), relevant previous
observational studies.
\citet{bamba2003, vink2003, yamazaki2004, ballet2006} consider advection and
diffusion separately (with the break-point at $\Ecut$ when advective and
diffusive length scales are equal;
\citet{parizot2006, araya2010} consider diffusion and advection simultaneously
with the catastrophic dump equation.
\citet{berezhko2003, berezhko2004, cassam-chenai2007, morlino2010, rettig2012}
consider diffusion and advection simultaneously with the continuous loss
equation instead.
Conclusion that magnetic fields are amplified according to loss-limited model.  How?  Field
amplification from plasma instabilities, evidence for stronger compression ratios?
\citep{blondin2001}.  Lots of numerical and theoretical work that I don't
understand.
\citet{rettig2012} mention independent estimates of magnetic field, from
\citet{acero2010, acciari2011} for SN 1006 and Tycho (considering gamma-ray,
radio, X-ray data).

Magnetic damping offers a tantalizing alternate explanation and has not
been fully tested.  Equivocal results by \citet{araya2010} for Cas A.
\citet{rettig2012} gave model predictions for several historical SNRs and
proposed discriminating based on filament spectra -- the
expectation is that damped spectra are softer, loss-limited harder.
\citet{ressler2014} developed the idea of investigating energy dependence and
found rim narrowing in the remnant of SN 1006, strongly favoring a loss-limited
model.

% Why Tycho's SNR
To further test these models, we turn to the remnant of Tycho's 1572 supernova
(hereafter, Tycho).  Tycho, being a nearby galactic SNR, is close enough that
its rims are well-resolved by \textit{Chandra}'s spatial resolution.  Tycho has
an extensive and unusually contiguous shell of synchrotron rims around its
periphery, consistent with expansion into a low density ISM; ISM.
\citet{williams2013} favor mean pre-shock ISM density $\abt 0.1$--$0.2
\unit{cm^{-3}}$, consistent with Tycho's X-ray expansion rate
\citep{katsuda2010}.  Although Tycho is expanding into an ISM gradient
\citep{williams2013} and interacting with a molecular cloud on its eastern limb
\citep{reynoso1999}, Tycho's spherical morphology suggests that we might expect
more consistent shock acceleration around its periphery than in other
historical remnants, whether of Type Ia or CC SNe \note[Aaron]{this is not really
justifiable, but I think it's worth noting something to this effect?}.
CR acceleration evidence in Tycho, if relevant? \citet{eriksen2011,
acciari2011}.

@@

% TODO Choose names for these models and refer to them consistently, throughout
% TODO define ALL symbols, synchrotron constants, etc (somewhere state that CGS
%      notation is being used)
% TODO double check equations, exposition against Rettig/Pohl and others
% TODO review variable definitions when closer to final draft
The energy and space distribution of electrons at a supernova remnant
controls the synchrotron rims we see in X-ray and radio.  We model electron
transport from initial injection at the remnant's forward shock as following
the advection-diffusion equation given by:
\note[Aaron]{ApJ prefers bold vectors and vector operators, fix later}
% TODO fix vector operators
\begin{equation}
  \frac{\ptl f}{\ptl t} + \del \cdot \left( f \vec{v} \right)
  = C + \del \cdot \left( D \del f \right)
\end{equation}
where $f = f(E,x)$ is electron distribution as a function of electron energy
$E$ and downstream distance $x$ (where the forward shock is at $x=0$ and
$x>0$ towards the remnant's interior); $C$ is an arbitrary source/sink term,
$D$ is the diffusion coefficient, and $\vec{v}$ is a velocity field.
By considering 1-D steady-state plane transport at a constant downstream
velocity $v_d$ \note[Aaron]{we threw out 1. spherical coordinate terms, and 2.
compressible flow terms, I think.  Is the latter justifiable?  Makes sense for
assumption of uniform downstream velocity -- but is this a safe assumption in
practice?  Double check order-of-magnitude...} and a space-independent diffusion
coefficient, we obtain:
\begin{equation}
    v_d \frac{\ptl f}{\ptl x} - D \frac{\ptl^2 f}{\ptl x^2} = C .
\end{equation}
The source/sink term $C$ is determined by our assumptions about the injected
electron energy spectrum and its time/space evolution.

@@

% ---------------------------
% FWHM measurement comparison
% ---------------------------
\section{Capped and background subtracted FWHM measurements, and simple model
fits}

@@

\begin{table}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filament 1.}
    \input{tab-fits-all-capsubbkg.tex}
\end{table}

\begin{table}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filaments 2--5.}
    \input{tab-fits-all-capsubbkg-pt2.tex}
\end{table}

\begin{table*}[ht]
    \scriptsize
    \centering
    \caption{Filament averages of best-fit parameters for constituent regions}
    \input{tab-fits-par-avg-capsubbkg.tex}
\end{table*}

@@

\begin{table}[ht]
    \tiny
    \centering
    \caption{Simple model best fits for individual regions, Filaments 1 and 2}
    \input{tab-fits-all-simp.tex}
\end{table}

\begin{table}[ht]
    \tiny
    \centering
    \caption{Simple model best fits for individual regions, Filaments 3--5}
    \input{tab-fits-all-simp-pt2.tex}
\end{table}

% --------------
% SN 1006 tables
% --------------
\section{Full model validation, SN 1006 (DRAFT ONLY)}

>>>>

% This paper
\newcommand*\mean[1]{\bar{#1}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand*{\tsup}{\textsuperscript}

\newcommand*{\Chandra}{\textit{Chandra}\ }
\newcommand*{\tsynch}{\tau_{\mt{synch}}}
\newcommand*{\mE}{m_\mt{E}}
\newcommand*{\Ecut}{E_{\mt{cut}}}
\defcitealias{ressler2014}{R14}

@@

% Intro -- shocks, DSA, thin rims, filament observations; other fields/work
Forward shock accelerated electrons in young supernova remnants (SNRs) emit
synchrotron radiation strongly in the shock's immediate wake but quickly turn
off downstream \citep{koyama1995, reynolds1996}, producing a shell-like
morphology with bright X-ray and radio rims/filaments due to line-of-sight
projection (hereafter, thin rims).  X-ray and gamma ray imaging has provided
several lines of indirect and direct evidence for hadron and lepton
acceleration in SNRs \citep{uchiyama2007, aharonian2004, acero2010,
ackermann2013}; electrons, in particular, are expected to be accelerated to
$\abt10$--$100\unit{TeV}$ energies.  Although synchrotron emission due to
accelerated electrons does not imply acceleration of an unseen hadronic
component, the prevailing explanatory mechanism of diffusive shock acceleration
(DSA) should operate on both protons and electrons; thus, observations of thin
synchrotron rims has been often taken as supporting, indirect evidence for
cosmic ray acceleration.  \note[Aaron]{first/second pass, very poor wording}

Spectral and spatial measurements of synchrotron rims can help constrain
physical parameters of the shock acceleration process.  Many previous workers
have inferred magnetic field amplification at shock fronts in SNRs; a host of
mechanisms \emph{what?!} may account for the thinness of the observed
shocks -- much thinner than $1/12$ the forward shock radius $r_s$ as would be
expected from a Sedov density profile \note[Aaron]{but Bamba et al. 2003 find
consistent scale length for SN 1006?}.  \note[Aaron]{need a reference -- I
don't have a sense of which mechanisms are most important, relevant, proven?!
even} Many plasma physics details -- Bell's cosmic ray current instability
\citep{bell2001} ties magnetic field amplification w/ maximum CR energy (push
towards and just above the PeV limit).  The processes observed in SNR shocks
are broadly relevant to many collisionless astrophysical shocks, such as
Earth's bow shock, stellar wind-blown bubbles, starburst galaxies
\citep{heckman1990}, jets of active galactic nuclei, galaxy clusters
\citep{van-weeren2010}, and even cosmological shocks \citep{miniati2000,
ryu2008}.

% What makes these shocks?
The widths of these X-ray rims are thought to be controlled by a combination of
synchrotron losses, particle transport, and magnetic fields immediately
downstream of the shock.  Synchrotron losses depend on the initial electron
energy distribution and the gradual decrease of electron energies downstream of
the shock.  High energy electrons near the shock efficiently radiate harder
synchrotron X-ray photons, but decline in energy as they are transported away
from the shock; this may be seen in spectral softening downstream of the shock
\citep[e.g.,][]{cassam-chenai2007}.  The transport process is driven by bulk
plasma advection downstream of the shock and collisionless particle diffusion
\note[Aaron]{src?! need a simple review of plasma diffusion.  What are the
relevant drivers of plasma diffusion?  Does it make sense to think of regimes
of particle interaction -- e.g., thinking of single particle random
interaction, vs. macroscopic MHD waves, or other structure?!} with respect to
bulk advection.  In general, diffusion allows higher energy electrons to
diffuse further upstream or downstream than would be expected from pure
advection.  Finally, the magnetic field strength may be quickly damped behind
the shock, so that electrons may not radiate as efficiently and thin rims would
reflect the spatial structure of the field, rather than efficient particle
acceleration (and losses) and/or efficient synchrotron cooling
\citep{pohl2005}.

% Energy dependence.
Moreover, these controlling mechanisms predict different scalings for rim width
as a function of energy.  Synchrotron loss-limited rims should scale as $w
\propto \nu^{-1/2}$.  Diffusion smears out rims at all energies with stronger
effect at higher energies, yielding a slower drop-off in rim widths when
diffusion dominates.  Magnetic damping, if at a length scale comparable to
filament widths, predicts comparatively energy-independent rim widths --
intuitively, if the magnetic field turns off, synchrotron radiation turns off
regardless of electron energy.  \note[aaron]{this would benefit from
numbers/quantification, I think this statement is stronger if the magnetic
field is stronger -- compare Figure 1 of \citet{pohl2005}}\citet{ressler2014}
give a much fuller exposition of the relative effects of advection, diffusion,
and magnetic field damping on filament rim widths.

% Which one is right?
Which mechanisms are most relevant to historical SNRs?

The primary route of attack has been to just measure the filament widths,
determine relevant advective/diffusive lengthscales, and perhaps favor one or
the other.  Assuming that thin rim widths are limited by synchrotron energy
losses (loss-limited rims), you have $l_{\mt{ad}} \propto v_d \tsynch$
gives a magnetic field...
(how did people look at $\Ecut$ or $D$?)
\citet{bamba2003, vink2003, yamazaki2004, ballet2006}
More sophisticated models have followed, but the basic result is consistent --
magnetic fields are amplified, with some amount of diffusion.
\citet{parizot2006, araya2010} consider diffusion and advection simultaneously
with the catastrophic dump equation.
\citet{berezhko2003, berezhko2004, cassam-chenai2007, morlino2010, rettig2012}
consider diffusion and advection simultaneously with the continuous loss
equation instead.

Some other approaches -- spectral softening \citep{cassam-chenai2007}, CD-FS
gap \citep{warren2005}, multi-wavelength observations \citep{acero2010,
acciari2011, morlino2012} for SN 1006 and Tycho.

Magnetic damping offers a tantalizing alternative explanation and has not
been fully tested.  Equivocal results by \citet{araya2010} for Cas A.
\citet{marcowith2010} did ???
\citet{rettig2012} gave model predictions for several historical SNRs and
proposed discriminating based on filament spectra -- the
expectation is that damped spectra are softer, loss-limited harder.
\citet{ressler2014} developed the idea of investigating energy dependence and
found rim narrowing in the remnant of SN 1006, strongly favoring a loss-limited
model.

% Why Tycho's SNR
To further test these models, we turn to the remnant of Tycho's 1572 supernova
(hereafter, Tycho).  Tycho, being a nearby galactic SNR, is close enough that
its rims are well-resolved by \textit{Chandra}'s spatial resolution.  Tycho has
an extensive and unusually contiguous shell of synchrotron rims around its
periphery, consistent with expansion into a low density ISM; ISM.
\citet{williams2013} favor mean pre-shock ISM density $\abt 0.1$--$0.2
\unit{cm^{-3}}$, consistent with Tycho's X-ray expansion rate
\citep{katsuda2010}.  Although Tycho is expanding into an ISM gradient
\citep{williams2013} and interacting with a molecular cloud on its eastern limb
\citep{reynoso1999}, Tycho's spherical morphology suggests that we might expect
more consistent shock acceleration around its periphery than in other
historical remnants, whether of Type Ia or CC SNe \note[Aaron]{this is not really
justifiable, but I think it's worth noting something to this effect?}.

@@

% TODO Choose names for these models and refer to them consistently, throughout
% TODO define ALL symbols, synchrotron constants, etc (somewhere state that CGS
%      notation is being used)
% TODO double check equations, exposition against Rettig/Pohl and others
% TODO review variable definitions when closer to final draft
The energy and space distribution of electrons at a supernova remnant
controls the synchrotron rims we see in X-ray and radio.  We model electron
transport from initial injection at the remnant's forward shock as following
the advection-diffusion equation given by:
\note[Aaron]{\citet{blandford1987} take a different and more thorough tack in
deriving the relevant transport equations, and do include the velocity
divergence term (start w/ Vlasov equation, then transform to non-inertial?
shock frame and tack on assumptions).  What is the best way to present this?
I'm not sure if na\"ively citing generalized ``transport'' is sufficient.}
% TODO fix vector operators
\begin{equation}
  \frac{\ptl f}{\ptl t} + \del \cdot \left( f \vec{v} \right)
  = C + \vec{\del} \cdot \left( D \vec{\del} f \right)
\end{equation}
where $f = f(E,x)$ is electron distribution as a function of electron energy
$E$ and downstream distance $x$ (where the forward shock is at $x=0$ and
$x>0$ towards the remnant's interior); $C$ is an arbitrary source/sink term,
$D$ is the diffusion coefficient, and $\vec{v}$ is a velocity field.
By considering 1-D steady-state plane transport at a constant downstream
velocity $v_d$ \note[Aaron]{we threw out 1. spherical coordinate terms, and 2.
compressible flow terms, I think.  Is the latter justifiable?  Makes sense for
assumption of uniform downstream velocity -- but is this a safe assumption in
practice?  Double check order-of-magnitude...} and a space-independent diffusion
coefficient, we obtain:
\begin{equation}
    v_d \frac{\ptl f}{\ptl x} - D \frac{\ptl^2 f}{\ptl x^2} = C .
\end{equation}
The source/sink term $C$ is determined by our assumptions about the injected
electron energy spectrum and its time/space evolution.

@@

% ---------------------------
% FWHM measurement comparison
% ---------------------------
\section{Additional calculations}

@@

I also present results for simple model fits, and fits for a larger set of
regions (22 instead of 13).

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Measured FWHMs for all regions, capped and background subtracted.}
    \input{tab-fwhms-capsubbkg.tex}
\end{table*}

\begin{table}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filament 1.  Capped
    and background subtracted FWHMs.}
    \input{tab-fits-all-capsubbkg.tex}
\end{table}

\begin{table}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filaments 2--5.
    Capped and background subtracted FWHMs.}
    \input{tab-fits-all-capsubbkg-pt2.tex}
\end{table}

\begin{table*}[ht]
    \scriptsize
    \centering
    \caption{Filament averages of best-fit parameters for constituent regions.
    Capped and background subtracted FWHMs.}
    \input{tab-fits-par-avg-capsubbkg.tex}
\end{table*}

@@

\begin{table}[ht]
    \tiny
    \centering
    \caption{Simple model best fits for individual regions, Filaments 1 and 2.
    Same FWHM as in main text.}
    \input{tab-fits-all-simp.tex}
\end{table}

\begin{table}[ht]
    \tiny
    \centering
    \caption{Simple model best fits for individual regions, Filaments 3--5.
    Same FWHM as in main text.}
    \input{tab-fits-all-simp-pt2.tex}
\end{table}

% -------------
% Region 5 data
% -------------

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Measured FWHMs for all regions (22).}
    \input{tab-fwhms-reg5.tex}
\end{table*}

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Full model best fits for smaller regions, filaments 1--3.  Same
    FWHM as in main text.}
    \input{tab-fits-all-reg5.tex}
\end{table*}

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Full model best fits for smaller regions, filaments 4--5.  Same
    FWHM as in main text.}
    \input{tab-fits-all-reg5-pt2.tex}
\end{table*}

\begin{table*}[ht]
    \scriptsize
    \centering
    \caption{Filament averages of best-fit parameters for constituent regions,
    with 22 instead of 13 regions.}
    \input{tab-fits-par-avg-reg5.tex}
\end{table*}

\begin{figure*}
    \centering
    \epsscale{0.7}
    \plotone{figures/f0-snr-reg5-inv.png}
    \caption{RGB image of Tycho with region selections overlaid.  Bands are
    0.7--1 keV (red), 1--2 keV (green) and 2--7 keV (blue).}
\end{figure*}

% --------------
% SN 1006 tables
% --------------
\clearpage
\section{Full model validation, SN 1006 (DRAFT ONLY)}
